Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-09-13 Paper/application.tex
polish application and intro

We evaluate a range of attack models for HTML, JavaScript, JSON, XML, and CSS. Listing \ref{lst-css-attack} shows the CSS attack model, which is used to verify the corresponding sanitizer. The attack model includes comments, declaration breakers, and known dangerous tokens such as \textbf{expression(}, \textbf{@import}, \textbf{behavior:}, \textbf{-moz-binding}, and \textbf{url(} (unless the URL is separately validated).
These are unsafe and undesired strings in CSS contexts.
Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and CleanseCSS, where "GA" means from Google Analytics. Sanitizers are often composed with preprocessing steps such as Trim, which removes leading and trailing whitespace characters. As shown in the table, GAHtmlEscape and escapeString are unsafe with respect to the HTML attack model without preprocessing, but become safe when combined with Trim. The \textbf{Size} column demonstrates the expressive power and succinctness of SFTs in modeling real-world transducers. We did not list the execution time of verification because all of them are quite efficent with less than 0.02 second to finish the verificaiton.

@@

  We can extend the above construction to model both replace-first and replace-all. In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}. The SFT $\mathcal{T}_2$ combines pattern matching and replacement using the construction just described. The SFT $\mathcal{T}_1$ matches strings that do not contain any substrings matching the pattern. Two $\varepsilon$-transitions connect these components: one from $\mathcal{T}_1$'s accepting state to $\mathcal{T}_2$'s initial state, and another from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state. This cyclic structure enables the replace-all operation by repeatedly applying pattern replacement until no more matches exist in the input string.

@@

Consider a string transformation that maps each character in $[\texttt{a-z}]$ to its successor, with \texttt{z} wrapping around to \texttt{a}. Modeling this transformation using only replacement operations requires spliting the interval $[\texttt{a-z}]$ into individual characters, resulting in 26 separate \texttt{replace-all} operations—one for each letter. For example:
\begin{verbatim}
  replace(replace(..., b, c, g), a, b, g)
\end{verbatim}

>>>>

We evaluate a range of attack models for HTML, JavaScript, JSON, XML, and CSS. Listing \ref{lst-css-attack} shows the CSS attack model, which is used to verify the corresponding sanitizer. The attack model includes comments, declaration breakers, and known dangerous tokens such as \textbf{expression(}, \textbf{@import}, \textbf{behavior:}, \textbf{-moz-binding}, and \textbf{url(} (unless the URL is separately validated).
These are unsafe and undesired strings in CSS contexts.
Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and CleanseCSS, where "GA" means from Google Analytics. Sanitizers are often composed with preprocessing steps such as Trim, which removes leading and trailing whitespace characters. As shown in the table, GAHtmlEscape and escapeString are unsafe with respect to the HTML attack model without preprocessing, but become safe when combined with Trim. The \textbf{Size} column demonstrates the succinctness of SFTs in modeling real-world transducers. We did not list the execution time of verification because all of them are quite efficent with less than 0.02 second to finish the verificaiton.

@@

  We can extend the above construction to model both replace-first and replace-all for XSS attack verification. In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}. The SFT $\mathcal{T}_2$ combines pattern matching and replacement using the construction just described. The SFT $\mathcal{T}_1$ matches strings that do not contain any substrings matching the pattern. Two $\varepsilon$-transitions connect these components: one from $\mathcal{T}_1$'s accepting state to $\mathcal{T}_2$'s initial state, and another from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state. This cyclic structure enables the replace-all operation by repeatedly applying pattern replacement until no more matches exist in the input string. Replace-first modeling has no transtion from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state.

@@

Consider a string transformation that replace each character in $[\texttt{a-z}]$ with its ASCII code. Modeling this transformation using only replacement operations requires spliting the interval $[\texttt{a-z}]$ into individual characters, resulting in 26 separate \texttt{replace-all} operations—one for each letter. For example:
\begin{verbatim}
  replace(replace(..., b, 0x62, g), a, 0x61, g)
\end{verbatim}

### 2025-09-13 Paper/introduction.tex
polish application and intro

Despite these, existing verified implementations for string analysis do not
handle the following indispensable string operations for analysis of web
applications (e.g. see
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}):
\emph{string transductions} and \emph{replaceall} operators. In particular, 
string transductions are used frequently in web applications (e.g. for
sanitization, as well as implicitly applied transductions, for instance
through \texttt{innerHTML}) and are well-known to be critical for analyzing XSS
vulnerabilities for web applications
\cite{Berkeley-JavaScript,systematic-transduction,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
arise from such applications require the \emph{replaceall} operator (part of
SMT-LIB), which is currently not handled by any existing verified
implementation. This paper tackles the formalization of SFA/SFT and
their associated algorithms, as well as applications to analysis of sanitizers
and verified solver for string constraints. 
%\textcolor{red}{Updated up to here
%so far}

>>>>

Despite these, existing verified implementations for string analysis do not
handle the following indispensable string operations for analysis of web
applications (e.g. see
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}):
\emph{string transductions} and \emph{replaceall} operators. In particular, 
string transductions are used frequently in web applications (e.g. for
sanitization, as well as implicitly applied transductions, for instance
through \texttt{innerHTML}) and are well-known to be critical for analyzing XSS
vulnerabilities for web applications
\cite{Berkeley-JavaScript,systematic-transduction,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
arise from such applications require the \emph{replaceall} operator, which is currently not handled by any existing verified
implementation. This paper tackles the formalization of SFA/SFT and
their associated algorithms, as well as applications to analysis of sanitizers
and verified solver for string constraints. 
%\textcolor{red}{Updated up to here
%so far}
